### 2018-04-16 contents/reviews/dmri.tex
Updates from ShareLaTeX

Diffusion magnetic resonance imaging (dMRI) is a popular MR modality that is capable of providing orientation information of brain tissue in-vivo, and thereby permitting the reconstruction and analysis of brain white matter structures with the use of computer reconstruction techniques.

@@

The basis of dMRI is based on the observation that the location probability distribution of water molecules, which undergoes Brownian motion during diffusion, is a function of its suspension medium. The rate of diffusion can be formulated as \cite{Mukherjee2008}: 
\begin{equation}
<r^2>=6Dt
\end{equation}
where $<r^2>$ is the mean square displacement of the molecule after time $t$ with diffusion constant $D$. In an unhindered medium, such as in liquid water, its diffusion probability is can be described as an spherical volume, in other words it undergoes \textit{isotropic} diffusion. In biological tissue such as the neuron axonal bundle, its diffusion probability is then constrained to be more likely in a direction along the bundle, rather than parallel to the cross-section, its diffusion state is then described as \textit{anisotropic} (\ref{fig:diffusion}).

@@

The image contrast is determined by the amount molecular motion between the two gradients, where motion along the gradient axis results in a loss of image intensity. The contrast is defined as: 
\begin{equation}
S_i = S_0 \cdot e^{-b \cdot ADC_i}
\end{equation}
Here $S_i$ denote the image contrast at a voxel. $S_0$ the signal intensity without the gradients applied. $ADC_i$ being the apparent diffusion coefficient in the $i$ direction. $b$ is the diffusion weighted factor that is a function is the duration, strength and temporal spacing of the gradients: 
\begin{equation}
b = \gamma^2 G^2 \delta^2(\Delta - \delta/3)
\end{equation}
where $\gamma$ is the gyromagnetic ratio constant. The ADC is measured in $mm^2/s$, and denotes the average molecular diffusion area with a given acquisition time. The higher ADC values indicate higher molecular mobility. Clinical DWI sequences typically use a diffusion time of 10-50 ms, or a travel of about 10$\mu$m. Typical clinical $b$ values range from 600 to 1500, to balance signal contrast with acquisition time.

@@

FA has been found to correlate with a wide spectrum of neuromorphological changes, including but not limited to multiple sclerosis \shortcite{Cercignani2001,Ciccarelli2003d,Miron2012}, vestibular schwannomas \shortcite{Chen2011b,Taoka2006,Wai2009}, tumors \shortcite{Mori2002,Schonberg2006,Nimsky2005}, schizophrenia \shortcite{Rametti2009,Fitzsimmons2009}, traumatic brain injury \shortcite{Bendlin2008,Kinnunen2011c,Hellyer2012}. RD is found to positively correlate with degree of demyelination in multiple sclerosis \shortcite{Song2005,Song2002,Klawiter2011,Janve2013}, and well as stages of white-matter development and aging \shortcite{Counsell2006,Sala2010}, including alzheimer's disease \shortcite{Acosta-Cabronero2012}. AD is correlated positively with axonal integrity \shortcite{Song2003}, and is found to be a correlate of axonal injury \shortcite{Budde2009} such as those found in traumatic brain injury \shortcite{Kinnunen2011c} and Gulf War illness \shortcite{Rayhan2013}. MD is correlated with tissue inflammation and edema \shortcite{Senda2012g}, and is found to correlate with diffusivity changes as a result of gliomas \shortcite{Stadlbauer2007}, multiple sclerosis \shortcite{Cercignani2001,Senda2012g,DeGroot2013}, amyotrophic lateral sclerosis \shortcite{Sharma2012}, and aging \shortcite{Kantarci2011}.

@@

Modifications to B-value needs to be considered for HARDI. The measured diffusion contrasts are dependent on the diffusion compartment variances. Therefore higher B-values will increase the measured signals variance contrasts, and allow better separation between crossing diffusion directions. Different HARDI algorithms have different requirements on B-values. For example Q-ball imaging needs B=3000 $s/mm^2$ \cite{Descoteaux2007a}, and multi-shell acquisitions will require stepped values of B values ranging from 500  $s/mm^2 $  to greater than 3500  $s/mm^2 $ \cite{Tournier2013}.

The determination of gradient sampling pattern and B-value directly affect acquisition time and image quality. More complex gradient sampling and higher B-values can improve tractography quality. However the acquisition time can become prohibitively long in a clinical setting. Other factors that must be considered include the nature of patient pathology, motion artifacts, and the inclusion of other imaging modalities during clinical scan sessions.

@@

Spherical decovolution is considered to be the action of a set of rotations on a function over a sphere. Given $F^l $ as a vector of $l$th order spherical harmonic decomposition of  $F(\theta,\phi)$, with vector length $(2l + 1)$, and $R^l$ as the $l$th order rotational harmonic decomposite of $R(\theta)$, with vector length $(2l+1)^2$. The spherical harmonic representation of $S(\theta,\phi)$ is then:

@@

The Tikhonov regularization minimizes the weighted sum with two terms:
\begin{equation}
\|{Af - b}^2\|^2 + \lambda^2 \| l(f - f')\|^2
\end{equation}
Where the first term is a linear least-square fit of solution $f$ to to data $b$ with problem matrix $A$. 
In terms of CSD, $f$ corresponds to the SH coefficients of the FOD, $b$ is the measured signal and $A$ is:
\begin{equation}
A = QR
\end{equation}
Where $Q$ is the SH coefficient to DW intensity mapping, and $R$ is the spherical deconvolution. 
The regularization term is the constrained matrix $L$ multiplying the norm of $f$ relative to an initial estimate $f'$. The parameter $\lambda$ controls the regularization amount, with $\lambda=0$ having no regularization, to $\lambda\rightarrow \infty $ making the solution tend to $f'$.

@@

\subsection{Tractography}
Tractography is defined as the construction, or propagation of streamline from an underlying field of orientation information, in order to visualize continuous structures within a volume.  The orientation field is generally provided as a 3D matrix of directional information obtained from DTI, CSD, or other methods. For example, in DTI the directions are the eigen vector with the greatest magnitude $\hat{\lambda_1}$; for CSD, the directional vectors would be peaks from the FOD.

Applying pathway tracking to DWI imaging was first explored by Mori \cite{Mori2002b} and Basser\cite{Basser2000}. The family of methods that propagates by deterministic algorithms based on local orientation data is collectively called deterministic tractography. The first tracking algorithms were applied to DTI using step-wise polyline propagators that followed local DT tensors \cite{Conturo1999}. Further refinement of the line propagator to include neighbouring DT tensors, and interpolation of the local DT directions in order to obtain a smooth curve \cite{Mori2002b}. Basser introduced Runge-Kutta integration to allow more robust path-tracing when faced with noise in the local directions \cite{Basser2000}. Another similar path-integral method was also proposed by Tuch \cite{Tuch2000d}. The integrated tractography approaches are now standard practice for deterministic tractography \cite{Tournier2004,Pieper2004}.

Deterministic tractography lends itself naturally to anatomical interpretation and further geometric analysis. It was traditionally paired with DTI, and so was thought to be limited in the case of crossing fibers. Other methods attempted to circumvent the limitation with probabilistic approaches that include white matter tissue priors in the fiber propagation model. The most popular of the probabilistic approach is the FSL software's probabilistic tractography module (FDT) \cite{Behrens2007}, which also includes a model-based ball-and-stick diffusion model that can resolve 2-fiber crossings. The FDT software produces visitation images of its probabilistic streams that collate the probability of tract propagation at each voxel.  The visitation image, however, is hard to interpret and localize, and therefore does not rival deterministic methods in surgical visualization.

@@

\subsubsection{Applications}
Tractography permits the mapping of white matter structures with computer graphic and statistic methods. It permits the study of brain structural connectivity at individual level \cite{Boorman2007} and group level \cite{Meskaldji2013},  circuitry mapping \cite{Kirsch2015} and macro-structural insights between the brain hemispheres \cite{Kucyi2012a}. It is often paired with functional connectivity for studies in multiple sclerosis \cite{Zhong2016,Rocca2007a,Kucyi2015}, chronic pain \cite{Mansour2013,Seifert2011,Wiech2014}, and other areas of study such as empathy \cite{Bernhardt2014}, and aphasia \cite{Catani2013c}.

@@

\begin{figure}[ht]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{tractography1}
\end{tabular}
\caption{Example of cranial nerve reconstruction with deterministic tractography.} 
\label{fig:tract1}
\end{center}
\end{figure}

>>>>

Diffusion magnetic resonance imaging (dMRI) is a popular MR modality that is capable of providing orientation information of brain white matter fibers \textit{in-vivo}, and thereby permitting the reconstruction and analysis of brain white matter structures with the use of computer reconstruction techniques.

@@

DMRI is based on the observation that the location probability distribution of water molecules, which undergoes Brownian motion during diffusion, is a function of its suspension medium. The rate of diffusion can be formulated as \cite{Mukherjee2008}: 
\begin{equation}
<r^2>=6Dt
\end{equation}
where $<r^2>$ is the mean square displacement of the molecule after time $t$ with diffusion constant $D$. In an unhindered medium, such as in liquid water, the water molecule diffusion probability is can be described as an spherical volume, in other words it undergoes \textit{isotropic} diffusion. In biological tissue such as the neuron axonal bundle, water molecule diffusion probability is then constrained to be more likely in a direction along the bundle, rather than parallel to the cross-section, its diffusion state is then described as \textit{anisotropic} (\ref{fig:diffusion}).

@@

The image contrast is determined by the amount molecular motion between the two diffusion gradients, where motion along the gradient axis results in a loss of signal intensity. The contrast is defined as: 
\begin{equation}
S_i = S_0 \cdot e^{-b \cdot ADC_i}
\end{equation}
Here $S_i$ denote the image contrast at a voxel. $S_0$ the signal intensity without the gradients applied. $ADC_i$ being the apparent diffusion coefficient in the $i$ direction. $b$ is the diffusion weighted factor that is a function of the duration, strength and temporal spacing of the gradients: 
\begin{equation}
b = \gamma^2 G^2 \delta^2(\Delta - \delta/3)
\end{equation}
where $\gamma$ is the gyromagnetic ratio constant. The ADC is measured in $mm^2/s$, and denotes the average molecular diffusion area with a given acquisition time. The higher ADC values indicate higher molecular mobility. Clinical DWI sequences typically use a diffusion time of 10-50 ms, or a travel of about 10$\mu$m. Typical clinical $b$ values range from 600 to 1500 $s/mm^2$, to balance signal contrast with acquisition time.

@@

Changes in white matter FA has been found to correlate with a wide spectrum of neuromorphological changes, including but not limited to multiple sclerosis \shortcite{Cercignani2001,Ciccarelli2003d,Miron2012}, vestibular schwannomas \shortcite{Chen2011b,Taoka2006,Wai2009}, tumors \shortcite{Mori2002,Schonberg2006,Nimsky2005}, schizophrenia \shortcite{Rametti2009,Fitzsimmons2009}, traumatic brain injury \shortcite{Bendlin2008,Kinnunen2011c,Hellyer2012}. RD is found to positively correlate with degree of demyelination in multiple sclerosis \shortcite{Song2005,Song2002,Klawiter2011,Janve2013}, and well as stages of white-matter development and aging \shortcite{Counsell2006,Sala2010}, including alzheimer's disease \shortcite{Acosta-Cabronero2012}. AD is correlated positively with axonal integrity \shortcite{Song2003}, and is found to be a correlate of axonal injury \shortcite{Budde2009} such as those found in traumatic brain injury \shortcite{Kinnunen2011c} and Gulf War illness \shortcite{Rayhan2013}. MD is correlated with tissue inflammation and edema \shortcite{Senda2012g}, and is found to correlate with diffusivity changes as a result of gliomas \shortcite{Stadlbauer2007}, multiple sclerosis \shortcite{Cercignani2001,Senda2012g,DeGroot2013}, amyotrophic lateral sclerosis \shortcite{Sharma2012}, and aging \shortcite{Kantarci2011}.

@@

Modifications to b-value needs to be considered for HARDI. The measured diffusion contrasts are dependent on the diffusion compartment variances. Therefore higher b-values will increase the measured signals variance contrasts, and allow better separation between crossing diffusion directions. Different HARDI algorithms have different requirements on b-values. For example Q-ball imaging needs b=3000 $s/mm^2$ \cite{Descoteaux2007a}, and multi-shell acquisitions will require stepped values of b-values ranging from 500  $s/mm^2 $  to greater than 3500  $s/mm^2 $ \cite{Tournier2013}.

The determination of gradient sampling pattern and b-value directly affect acquisition time and image quality. More complex gradient sampling and higher b-values can improve tractography quality. However the acquisition time can become prohibitively long in a clinical setting. Other factors that must be considered include the nature of patient pathology, motion artifacts, and the inclusion of other imaging modalities during clinical scan sessions.

@@

Spherical covolution is considered to be the action of a set of rotations on a function over a sphere. Given $F^l $ as a vector of $l$th order spherical harmonic decomposition of  $F(\theta,\phi)$, with vector length $(2l + 1)$, and $R^l$ as the $l$th order rotational harmonic decomposite of $R(\theta)$, with matrix size $(2l+1)^2$. The spherical harmonic representation of $S(\theta,\phi)$ is then:

@@

The Tikhonov regularization minimizes the weighted sum with two terms:
\begin{equation}
\|{Af - b}\|^2 + \lambda^2 \| l(f - f')\|^2
\end{equation}
Where the first term is a linear least-square fit of solution $f$ to to data $b$ with problem matrix $A$. 
In terms of CSD, $f$ corresponds to the SH coefficients of the FOD, $b$ is the measured signal and $A$ is:
\begin{equation}
A = QR
\end{equation}
Where $Q$ is the SH coefficient to diffusion weighted intensity mapping, and $R$ is the spherical deconvolution. 
The regularization term is the constrained matrix $L$ multiplying the norm of $f$ relative to an initial estimate $f'$. The parameter $\lambda$ controls the regularization amount, with $\lambda=0$ having no regularization, to $\lambda\rightarrow \infty $ making the solution tend to $f'$.

@@

\subsection{Tractography}

Tractography is a computer graphics technique used by neuroscientists to delineate neural pathways with DWI imaging. Applying pathway tracking to DWI imaging in the brain was first explored by Mori \cite{Mori2002b} and Basser\cite{Basser2000}. Tractography in essence is the construction, or propagation of streamline from an underlying field of orientation information, in order to visualize continuous structures within a volume.  The orientation field is generally provided as a 3D matrix of directional information obtained from DTI, CSD, or other methods. For example, in DTI the directions are the eigen vector with the greatest magnitude $\hat{\lambda_1}$; for CSD, the directional vectors would be peaks from the FOD.

The family of methods that propagates by deterministic algorithms based on local orientation data is collectively called deterministic tractography. The first tracking algorithms were applied to DTI using step-wise polyline propagators that followed local DT tensors \cite{Conturo1999}. Further refinement of the line propagator to include neighbouring DT tensors, and interpolation of the local DT directions in order to obtain a smooth curve \cite{Mori2002b}. Basser introduced Runge-Kutta integration to allow more robust path-tracing when faced with noise in the local directions \cite{Basser2000}. Another similar path-integral method was also proposed by Tuch \cite{Tuch2000d}. The integrated tractography approaches are now standard practice for deterministic tractography \cite{Tournier2004,Pieper2004}. Deterministic tractography lends itself naturally to anatomical interpretation and further geometric analysis. It was traditionally paired with DTI, and so was thought to be limited in the case of crossing fibers.

Other methods attempted to circumvent the limitation with probabilistic approaches that include white matter tissue priors in the fiber propagation model. The most popular of the probabilistic approach is the FSL software's probabilistic tractography module (FDT) \cite{Behrens2007}, which also includes a model-based ball-and-stick diffusion model that can resolve 2-fiber crossings. The FDT software produces visitation images of its probabilistic streams that collate the probability of tract propagation at each voxel. The visitation image, however, is hard to interpret and localize, and therefore does not rival deterministic methods in surgical visualization.

@@

\subsubsection{Applications}
Tractography permits the mapping of white matter pathway structures with computer graphic and statistic methods. It permits the study of brain structural connectivity at individual level \cite{Boorman2007} and group level \cite{Meskaldji2013},  circuitry mapping \cite{Kirsch2015} and macro-structural insights between the brain hemispheres \cite{Kucyi2012a}. It is often paired with functional connectivity for studies in multiple sclerosis \cite{Zhong2016,Rocca2007a}, chronic pain \cite{Mansour2013,Seifert2011,Wiech2014}, and other areas of study such as empathy \cite{Bernhardt2014}, and aphasia \cite{Catani2013c}.

@@

\begin{figure}[ht]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{tractography1}
\end{tabular}
\caption{Example of cranial nerve reconstruction with deterministic tractography. The image represents a saggital view of a vestibular schwannoma (green) displacing the trigeminal nerve (blue) and the facial/vestibular nerve complex (red) at the level of the brainstem. } 
\label{fig:tract1}
\end{center}
\end{figure}

### 2018-04-16 contents/reviews/machine-learning.tex
Updates from ShareLaTeX

\section{Machine Learning}
Machine learning (ML) is a branch of computer science that aims to learn and form predictions from data. In contrast with static algorithmic programming, where the structures of the algorithms are predefined, machine learning approaches construct models based on observed data, in order to extend its predictions to the unobserved. Strategies and algorithms for constructing various models overlap with the fields of artificial intelligence, computational statistics, pattern recognition, and neuroscience. Advanced in modern machine learning, especially in deep neural networks \cite{LeCun2015}, in recent years have led to wide spread adaptation of machine learning based online services in translation \cite{Bahdanau2014a}, image search \cite{simonyan2014very}, and speech recognition \cite{graves2013speech}. There is also surging interest in deploy machine learning systems in radiology and disease diagnosis in medicine\cite{Kamnitsas2016,Suk2014}.

Machine learning training strategies can be categorized into Supervised Learning, Unsupervised Learning, and Reinforcement Learning. Supervised learning involves training data that are pre-labelled. For example, the data supplied can be a list of patients' age and gender, and the label can be whether they or not they are in pain. Unsupervised learning involves data without any prior labels, and the algorithm attempts to discover patterns embedded within, and is often used in data mining. Reinforcement learning involves continuously learning with a reward function. It is generally used to train agents that have a continuous stream of data and needs to interact with an external environment and is widely used in robotics.

@@

There is an extensive history of applying statistical methods to the field of neuroimaging, from the very onset of the field. Neuroimaging studies often attempt to correlate physiological and image-based measures with neuro-activities with specific experimental variables or conditions. Neuro-activities may be measured from EGG sensors, or as voxels in functional or structural MR images. Traditionally data from each measured feature component was assumed to be independent. These components maybe individual sensor readings, or voxel-based measures. Correlative analysis were performed using general linear models, with post-hoc false-positive corrections. This type of univariate analysis do not have the ability to discover key relationships between feature components. Therefore univariate models cannot take into account spatial or temporally correlative dynamics.

Increasing complexity in neuroimaging data has lead to widespread applications of multi-variable machine learning methods in neuroimaging. Application of ML in pain neuroimaging became controversial when Tor Wager \cite{Wager2013} demonstrated a Lasso-PCA model that classifies nociceptive pain with sensitivity and specificity of 90\% solely from fMRI imaging data. Additional studies showed consistently good performance of ML-based nociceptive pain discrimination. Discrimination of chronic pain from nociceptive pain proved to be more challenging, due to the high degree of similarity in pain-matrix activation patterns. The improvement in chronic pain detection lead to debates of the validity of presenting such methods as evidences in courtroom, where Davis et al argued against such actions \cite{Davis2012a}. While it is commonly accepted that pain activity in the brain involves multiple regions of the pain-matrix, evidences in some studies suggested that the posterior-insula is heavily involved in chronic pain. Segerdale et al claimed that the posterior insula is a primary pain region \cite{Segerdahl2015a}, thereby ignited renewed debates surrounding regional specific brain centers for chronic pain \cite{Davis2015}.

Despite the controversies, interest in machine learning in medical imaging prognosis have sparked both public and private efforts across the globe. Classification studies with using multi-modal data with features in functional connectivity and structural connectivity in pain syndromes such as trigeminal neuralgia continue to improve. It is clear that machine learning is here to stay in pain neuroimaging research.

>>>>

\section{Machine Learning}
Machine learning (ML) is a branch of computer science that aims to learn and form predictions from data. In contrast with static algorithmic programming, where the structures of the algorithms are predefined, machine learning approaches construct models based on observed data, in order to extend its predictions to the unobserved. Strategies and algorithms for constructing various models overlap with the fields of artificial intelligence, computational statistics, pattern recognition, and neuroscience. Advances in modern machine learning, especially in deep neural networks \cite{LeCun2015}, in recent years have led to wide-spread adoption of machine learning based online services in translation \cite{Bahdanau2014a}, image search \cite{simonyan2014very}, and speech recognition \cite{graves2013speech}. There is also surging interest in applying machine learning systems in medicine, particularly in radiology to assist in and disease diagnosis \cite{Kamnitsas2016,Suk2014}.

Machine learning training strategies can be categorized into Supervised Learning, Unsupervised Learning, and Reinforcement Learning. Supervised learning uses data that are pre-labelled according to some measure, in order to train a model. For example, the data supplied can be a list of patients' age and gender, and the label can be whether or not they are in pain. Unsupervised learning is a strategy that uses data without any prior labels, and the algorithm attempts to discover patterns embedded. Thus this approach is often used in data mining. Reinforcement learning involves continuously learning with a reward function. It is generally used to train autonomous agents using continuous streams of data, and is widely used in robotics.

@@

There is an extensive history of applying statistical methods to the field of neuroimaging, from the very onset of the field. Neuroimaging studies often attempt to correlate physiological and image-based measures with neuro-activities with specific experimental variables or conditions. Neuro-activities may be measured from EEG sensors, or as voxels in functional or structural MR images. Traditionally, data from each measured feature component was assumed to be independent. These components may be individual sensor readings or voxel-based measures. Correlative analysis were performed using general linear models, with post-hoc false-positive corrections. This type of univariate analysis does not have the ability to discover key relationships between feature components. Therefore univariate models cannot take into account spatial or temporally correlative dynamics.

Increasing complexity in neuroimaging data has led to widespread applications of multi-variable machine learning methods in neuroimaging. Application of ML in pain neuroimaging became controversial when Tor Wager \cite{Wager2013} demonstrated a Lasso-PCA model that classified acute experimental pain with sensitivity and specificity of up to 90\% solely from fMRI imaging data. Additional studies showed consistently good performance of ML-based nociceptive pain discrimination. Discrimination of chronic pain from nociceptive pain proved to be more challenging, due to the high degree of similarity in activation patterns. The development of pain detection through brain imaging led to debates of the validity of presenting such methods as evidences in courtroom, where Davis et al argued against such actions \cite{Davis2012a}. While it is commonly accepted that pain evokes activity in the brain in multiple regions of the dynamic pain connectome, some studies suggest that the dorsal posterior-insula is heavily involved in chronic pain. Segerdale et al claimed that the dorsal posterior insula is a primary pain region \cite{Segerdahl2015a}, and this proposition ignited renewed debates surrounding regional specific brain centers for chronic pain \cite{Davis2015}.

Despite the controversies, interest in machine learning in medical imaging prognosis have sparked both public and private efforts across the globe. Classification studies using multi-modal data with features involving functional connectivity and structural connectivity in pain syndromes such as TN continue to improve. It is clear that machine learning is here to stay in pain neuroimaging research.

### 2018-04-16 contents/reviews/mri.tex
Updates from ShareLaTeX

Protons aligned in the parallel alignment are more preferred as it has a lower energy state than in anti-parallel orientation. The energy E of a magnetic moment $\vec{\mu}$ when in a magnetic field $B_0$ is given by: $E=-\vec{\mu}\cdot \textbf{B}_0=-\mu_z B_0=-\gamma m \hbar B_0$. The lower energy state of protons corresponding to m = +1/2 in a static magnetic field $B_0$ is $E(\frac{1}{2})=-\frac{1}{2}\gamma \hbar B_0$. The higher energy state of protons corresponding to m =-1/2 is given by $E(-\frac{1}{2})=\frac{1}{2}\gamma \hbar B_0$.
Consequently, the energy differences between two quantum states can be calculated as $\bigtriangleup E = \gamma \hbar B_0$ (Figure \ref{fig:energy}).

@@

Immediate after excitation, the phase precessions are synchronous, and therefore are in \textit{phase coherence}.  
However due to various reasons, the various spins begin to lose coherence, and their total sum magnetization $M_xy$ diminishes.

@@

\item There are intrinsic inhomogeneities by the magnetic field generator, and the scanning subject. These intrinsic properties contribute to the dephasing, and has a shorter time decay than T2. This type of decay time constant is named T2*  (Figure \ref{fig:T1T2}(C)). The loss of the MR signal due to T2* is called\textit{ free induction decay} (FID). T2* is defined as follows: 
$$ \frac{1}{T_2^*} = \frac{1}{T_2} + \frac{1}{T_2inhomo}$$ 
\end{enumerate}

@@

MRI signal intensity depends primarily on three parameters: 
\begin{itemize}
\item Proton density: the number of excitable spins per unit volume, which determines the maximum amount of signal that can be measured from the tissue. It can be measured by minimizing T1 and T2 times. 
\item T1 time: the amount time for $M_xy$ to recover to $M_z$, to be ready for re-excitation.
\item T2 time: the amount of time for $M_xy$ signal to fade under decoherence. 
\end{itemize}
The contrast in brain tissues of the resulting MR image can be tuned by adjusting these three parameters in pulse sequence design.

@@

\paragraph{Echo Time}
MR sequences have a gradient field which is used to control field inhomogeneities and to encode spatial origin. The gradients also contribute to spin dephasing. Therefore a refocusing pulse is necessary to measure MR signals adequately. The time delay between refocusing and the measured signal under phase coherence (spin echo) is known as echo time (TE). When TE is short ($<$ 30 ms), there are little signal differences between tissue types; when TE is long ( $\geq$ 60 ms ), the intrinsic tissue differences will allow good contrast in the resulting image. Therefore long TE is said to be T2 weighted.

\begin{figure}[htb]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{mri/spin-echo.pdf}
\end{tabular}
\caption{Illustration of the process of spin-echo. Initially, the spin precesses about the B0 field under steady-state. When a 90deg RF pulse is applied, the spins orients towards the xy-plane. When the RF pulse is switched off, the spins now relaxes towards the z-axis, and the spin undergoes phase decoherence. A 180 degrees flip angle pulse is applied to prevent spin saturation, and then a refocusing pulse is applied to bring spins back into coherence.  } \label{fig:spin-echo}
\end{center}
\end{figure}

Typically, T1-weighted spin echo (SE)  sequence is acquired with TR/TE at 340/13 ms; T2-weighted fast spin echo (FSE) at TR/TE of 3500/120 ms. 
In the case of T1-weighted MRI, images are created with short TE and short TR times. They can differentiate fat from water - with fat brighter and water darker (\url{www.mr-tip.com}) by using a gradient echo (GRE) sequence. In the brain T1-weighted scans provide good gray matter/white matter contrast. In addition, the image contrast can be increased with the use of an inversion pulse as in an MP-RAGE sequence. Due to short TR, this scan can be run very quickly, allowing the collection of high resolution 3D structural image (Figure \ref{fig:T1T2PD}, Left). These structural images can provide detailed information on the brain's shape and size, and provide the base for cortical surface reconstruction and segmentation.

The T2-weighted sequence uses a long TR and long TE. The T2-weighted sequence can be employed as a dual echo sequence. Like the T1-weighted scan, fat is
differentiated from water, but with fat dark and water bright (Figure \ref{fig:T1T2PD}, Middle).

The first or shorter echo ($T_E<30msec$) is proton density (PD) weighted or a mixture of T1 and T2. This image can provide good contrast of tissue and CSF, thus very helpful for evaluating periventricular pathology (Figure \ref{fig:T1T2PD}, Right). $T_2^*$-weighted scans use a gradient echo (GRE) sequence, with long TE and long TR. It is prone to susceptibility losses at air/tissue boundaries, but can increase contrast for certain types of tissue, such as venous blood. It is used for functional MRI detection allowing people to visualize and map the parts of the brain used to perform tasks.

@@

\begin{itemize}
\item Phase encoding: The phase encoding gradient is applied in the y-direction. It alters the Larmor frequency of the spins based the location of the y-axis, and results in phase shifts of the spins relative to each other. The phase shift is determined by the duration and amplitude of the gradient, when the gradient is switched off, all spins return to their initial rate of precession, but maintains the relative phase shift. 
\item Frequency encoding: The frequency encoding gradient is applied on the x-direction. This applies a magnetic field that increases in strength from right to left. The change in Larmor frequency alters the precession speed of the spins such that the spins towards the right precess faster than that of the left. The MR signal collected represents the entire frequency spectrum of the spins. 
\end{itemize}

@@

The collected MR data are encoded in a mathematical space termed the k-space. The k-space is a two-dimensional feature space, where the x-axis $k_x$ represent the measured frequency domain, and $k_y$ the phase domain. Each step in $k_y$ represents a single phase encoding step. The x-axis origin $k_x=0$ is the gradient iso-center, and the frequency distribution is symmetric about the x-origin. The 2D-FT operation is applied to the k-space to transform the frequency domain information into an image.

>>>>

Protons aligned in the parallel alignment are more preferred as they have a lower energy state than protons in anti-parallel orientation. The energy E of a magnetic moment $\vec{\mu}$ when in a magnetic field $B_0$ is given by: $E=-\vec{\mu}\cdot \textbf{B}_0=-\mu_z B_0=-\gamma m \hbar B_0$. The lower energy state of protons corresponding to m = +1/2 in a static magnetic field $B_0$ is $E(\frac{1}{2})=-\frac{1}{2}\gamma \hbar B_0$. The higher energy state of protons corresponding to m =-1/2 is given by $E(-\frac{1}{2})=\frac{1}{2}\gamma \hbar B_0$.
Consequently, the energy differences between two quantum states can be calculated as $\bigtriangleup E = \gamma \hbar B_0$ (Figure \ref{fig:energy}).

@@

Immediate after excitation, the phase precessions are synchronous, and therefore are in \textit{phase coherence}.  
However due to the following reasons, the various spins begin to lose coherence, and their total sum magnetization $M_xy$ diminishes.

@@

\item There are intrinsic inhomogeneities by the main static magnetic field generator, and the scanning subject. These intrinsic properties contribute to the dephasing, and has a shorter time decay than T2. This type of decay time constant is named T2*  (Figure \ref{fig:T1T2}(C)). The loss of the MR signal due to T2* is called\textit{ free induction decay} (FID). T2* is defined as follows: 
$$ \frac{1}{T_2^*} = \frac{1}{T_2} + \frac{1}{T_2inhomo}$$ 
\end{enumerate}

@@

MRI signal intensity depends primarily on three parameters: 
\begin{itemize}
\item Proton density: the number of excitable spins per unit volume, which determines the maximum amount of signal that can be measured from the tissue. It can be measured by minimizing T1 and maximizing T2 times. 
\item T1 time: the amount time for $M_xy$ to recover to $M_z$, to be ready for re-excitation.
\item T2 time: the amount of time for $M_xy$ signal to fade under decoherence. 
\end{itemize}
The contrast in brain tissues of the resulting MR image can be tuned by adjusting T1  and T2 parameters in pulse sequence design.

@@

\paragraph{Echo Time}
MR sequences have a gradient field which is used to encode spatial origin. The gradients also contribute to spin dephasing. In spin echo imaging, a 90 deg pulse and a 180 deg refocusing pulse are needed to measure MR signals adequately. The time delay between the first pulse and the measured signal under phase coherence (spin echo) is known as echo time (TE). When TE is short ($<$ 30 ms), and TR is short, the image is said to be T1 weighted; when TE is long ( $\geq$ 60 ms ), and TR is long, the image is said to be T2-weighted.

\begin{figure}[htb]
\begin{center}
\begin{tabular}{c}
\includegraphics[width=\textwidth]{mri/spin-echo.pdf}
\end{tabular}
\caption{Illustration of the process of spin-echo. Initially, the spins orients to wards B0 field under steady-state. When a 90deg RF pulse is applied, the spins orients to the xy-plane. When the RF pulse is switched off, the spins undergoe phase decoherence. A 180 degrees refocusing pulse is applied to bring spins back into coherence, at which point in time the peak of an "echo" occurs.  } \label{fig:spin-echo}
\end{center}
\end{figure}

Typically, T1-weighted spin echo (SE)  sequence is acquired with TR \textless 800 ms, TE \textless 20 ms; T2-weighted fast spin echo (FSE) at TR \textgreater 2000 ms, TE \textgreater 60 ms. 
In the case of T1-weighted MRI, images are created with short TE and short TR times. They can differentiate fat from water - with fat brighter and water darker (\url{www.mr-tip.com}) by using a gradient echo (GRE) sequence. In the brain T1-weighted scans provide good gray matter/white matter contrast. In addition, the image contrast can be increased with the use of an inversion pulse as in an MP-RAGE sequence. Due to short TR, this scan can be run very quickly, allowing the collection of high resolution 3D structural image (Figure \ref{fig:T1T2PD}, Left). These structural images can provide detailed information on the brain's shape and size, and provide the base for cortical surface reconstruction and segmentation.

The T2-weighted sequence uses a long TR and long TE. The T2-weighted sequence can be employed as a multi-echo sequence. Like the T1-weighted scan, fat is
differentiated from water in single echo sequences, but with fat dark and water bright (Figure \ref{fig:T1T2PD}, Middle).

Another type of image with shorter echo ($T_E<30msec$) is proton density (PD) (Figure \ref{fig:T1T2PD}, right) weighted. This image provides poor contrast of tissue and CSF, but good bright contrast against lesions, thus very helpful for evaluating periventricular pathology . $T_2^*$-weighted scans use a gradient echo (GRE) sequence, which uses only one RF pulse and gradient reversal. It is prone to susceptibility losses at air/tissue boundaries, but can increase contrast for certain types of tissue, such as venous blood. It is used for functional MRI detection allowing people to visualize and map the parts of the brain used to perform tasks.

@@

\begin{itemize}
\item Phase encoding: The phase encoding gradient is applied in the y-direction. It temporarily alters the Larmor frequency of the spins based their location of the y-axis, and results in phase shifts of the spins relative to each other. The phase shift is determined by the duration and amplitude of the gradient, when the gradient is switched off, all spins return to their initial rate of precession, but maintains the relative phase shift. 
\item Frequency encoding: The frequency encoding gradient is applied on the x-direction. This applies a magnetic field that increases in strength from right to left. The change in Larmor frequency alters the precession speed of the spins such that the spins towards the right precess faster than that of the left. The MR signal collected represents the entire frequency spectrum of the spins. 
\end{itemize}

@@

The collected MR data are encoded in a mathematical space termed the k-space. The k-space is a two-dimensional feature space, where the x-axis $k_x$ represent the measured frequency domain, and $k_y$ the phase domain. Each step in $k_y$ represents a single phase encoding step. The x-axis origin $k_x=0$ is the gradient iso-center, where the gradient magnetic field equals to zero, and the frequency distribution is symmetric about the x-origin. The 2D-FT operation is applied to the k-space to transform the frequency domain information into an image.
